---
title: "Homework 03"
author: "Helena Parsons"
format: pdf
---

# Part 1: Setting up tasks

## Reading in packages

```{r}
#| label: reading-in-packages
#| message: false
#| warning: false
library(tidyverse)
library(here)
library(janitor)
library(readxl)
```

## Creating data object

```{r}
#| label: creating-kelp-object
#| message: false
#| warning: false

# creating new object for temp-kelp data
kelp <- 
  # reading csv
  read_csv(
    # directing function to data file 
    here("data", "temp-kelp.csv"))

# creating new object for personal data 
job_search <- 
  # reading csv
  read_csv(
    # directing function to data file
    here("data", "personal-data.csv"))
```

# Part 2: Problems

## Problem 1: Giant kelp fronds

***Climate change threatens all ecosystems, including giant kelp (Macrocystis pyrifera) forests.***

***You are interested in the potential relationship between ocean temperature (measured in ° C) and giant kelp frond elongation rate (also known as growth rate, measured in cm day-1).***

***You conduct a field study in which you track giant kelp individuals growing in different temperatures (on average) and measure their elongation rates. Admittedly, this is not a perfect study, but you do the best with what you have!***

### 1a. An appropriate test

***In 1-3 sentences, name the appropriate test(s) to determine the strength of the relationship between temperature and giant kelp frond elongation rate (hint: there are two). Describe the differences between the two tests.***

***Be specific in your response to demonstrate your understanding of the variables in this question.***

Because I solely want to understand the strength of the relationship between temperature and giant kelp frond elongation rate, I want to use a correlation test.

A Pearson's r is used when there is a linear relationship between two variables. A Spearman's rho is used when the relationship between two variables is monotonic.

### 1b. Create a visualization

***Create a visualization that would be appropriate for showing the relationship between temperature and giant kelp frond elongation rate.***

***In addition to using the correct geometries, be sure to:***

-   ***relabel the x- and y-axes and include units***
-   ***use different colors from the ggplot() defaults***
-   ***use a different theme from the ggplot() default***

```{r}
#| label: visualiziation-of-temp-kelp

# creating a new object
temp_kelp_visual <- 
  # base layer: ggplot
  ggplot(
    # df: kelp
    data = kelp,
    # x-axis: temp_c
    # y-axis: kelp_elong
       mapping = aes(x = temp_c,
                     y = kelp_elong)) +
  # first layer: scatterplot
  geom_point(
    # changing color of points
    color = "magenta4") +
  # changing theme
  theme_classic() +
  # changing title and axes titles
  labs(title = "Kelp frond elongation rates decrease as temp increases",
       x = expression("Temperature ("*~degree*C*")"),
       y = expression("Giant kelp frond elongation rates (cm/day^-1)"))

# displaying object
temp_kelp_visual
```

### 1c. Check your assumptions and run your test

#### Assumptions

To run a Pearson's r, you want to ensure that there is a linear relationship between the two variables. From the scatter plot I created in part 1b, I can conclude that the relationship mostly appears to be linear.

```{r}
#| label: check-for-linear-relationship

# displaying object from 1b
temp_kelp_visual
```

I also want to make sure that the data is normally distributed. For this, I can create a QQ plot.

```{r}
# creating new object
qqplot_temp <- 
  # baselayer: ggplot
  ggplot(
    # df: kelp
    data = kelp,
    # variable: temp
       aes(sample = temp_c)) +
  # first layer: distribution of temp observations
       stat_qq(distribution=qnorm) +
  # second layer: expected distribution of temp
stat_qq_line( col = "black")

# displaying object
qqplot_temp
```

```{r}
# creating new object
qqplot_kelp <- 
  # first layer: ggplot
  ggplot(
    # df: kelp
    data = kelp,
         # variable: kelp_elong
       aes(sample = kelp_elong)) +
  # first layer: distribution of kelp elongation rate observations
  stat_qq(distribution=qnorm) +
  # second layer: expected distribution of kelp elongation rate
  stat_qq_line( col = "black")

# displaying object
qqplot_kelp
```

It seems that the variables are mostly normally distributed.

Additionally, the variables of kelp frond elongation rates and temperature need to be continuous, which they are since they are not integers, categorical, or ordinal data.

Finally, the observations also need to be independent, which is true because the observations of kelp growth rates do not all come from the same individual.

All of our assumptions have been checked so we can now run the Pearson's r.

#### Pearson's r test

```{r}
#| label: pearsons-r

#correlation test
cor.test(
  # df: kelp
  # first variable: temp
  kelp$temp_c, 
  # df: kelp
  # second variable: kelp elongation rates
         kelp$kelp_elong,
  # type of correlation test: pearson
         method = "pearson")
```

### 1d. Results communication

***In 1-3 sentences each, write about:***

-   ***Which test you used, and why (i.e. “To evaluate the strength of the relationship between temperature and giant kelp frond elongation rate, I used a…”)***
-   ***Your interpretation of your test (along with the appropriate summary of the test in parentheses)***

To analyze the strength of the relationship between temperature and giant kelp frond elongation rate, I used a Pearson's r correlation test.

I found that there is a moderate negative relationship between temperature and giant kelp frond elongation rate (Pearson's r = -0.69, t(30) = -5.2, p \< 0.001, $\alpha$ = 0.05).

### 1e. Test implications

***You’re working on a team of people who are also interested in the potential relationships between temperature and giant kelp growth rate, and more broadly giant kelp health. In 2-3 sentences, write what you would communicate to them about the results of this test and what it means for giant kelp growth.***

***Be cognizant of your audience as you are writing: what would they need to know to take action?***

I found a moderate negative (Pearson's r = -0.69) relationship between temperature and giant kelp frond elongation rate. Elongation rates are highest when temperature is below 16 °C and begins to decrease from there. While other factors may influence the overall health of giant kelp, the moderate correlation between temperature and frond elongation rates suggests that understanding changes in temperature and overall climate will be a significant factor in monitoring giant kelp growth.

